Mechanized Reasoning about a Capability Machine

### 2019-10-18 prisc-2020/paper.tex
Some editing
- the PC is not incremented when we are jumping, so I replaced with "updated" instead

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Capability machines are promising targets for secure compilers since capabilities can be used to enforce abstractions that are usually expected for high-level languages, such as well-bracketed control-flow (WBCF) and local state encapsulation (LSE).
  We present the first formalization of a capability machine that supports mechanized reasoning about deep semantic properties, including WBCF and LSE.
  Our formalization is done in the Coq implementation of Iris, a state-of-the-art concurrent higher-order separation logic, and includes a formalization of the logical relation defined by Skorstensgaard et. al. \cite{skorstengaardESOP18}, which can used to prove WBCF and LSE.
\end{abstract}

@@

While Iris is a framework and supports many languages, it is geared towards models of higher-level languages, which abstract from the fact that programs are stored in memory, and hence come equipped with notions of expressions and values, in addition to the program memory.
In contrast, our low-level capability machine model has no notion of expression and values, it just consists of a memory and a register file.
The program counter register contains a pointer to an address in memory.
That address in turn will contain an integer, which can then be decoded to a machine instruction, such as Load, Store, Jump, etc.
Once an instruction has been executed, the program counter is incremented and will then point to the next instruction in memory.

@@

% Similar to \cite{Jensen:2013}, the program logic rules for each instruction are of the following form: 
The Hoare triples for basic instructions take the following form
\begin{align*}
\small
	\textit{run time conditions } \wedge &\textit{ decode}(w) = instr \Rightarrow  \\
	\{\{\{ \textsf{PC} \mapsto_r ((p,g),b,e,a) &* a \mapsto_a[p] w * ... \}\}\} \\ 
	\textsf{Instr Executable }\\
	\{\{\{\textsf{PC} \mapsto_r ((p,g),b,e,a&+1) * a \mapsto_a[p] w * ... \}\}\}
\end{align*} 
Here, the run time conditions correspond to the dynamic checks done by the capability machine, and \textsf{Instr Executable} is the abstract expression for executing the next instruction in memory.
This form of Hoare triple is similar to the one used in \cite{Jensen:2013} but
unlike \cite{Jensen:2013}, the decoding function is in our case assumed.

@@

\newcommand{\World}{\mathrm{World}} \newcommand{\Word}{\mathrm{Word}} We now outline how we define a unary logical relation in Iris that captures capability safety.
We define a value relation $\mathcal{V}$ as an Iris relation of type $\World \to \Word \to iProp$, where $\World$ is a collection of state transition systems used to reason about local state.
(It is well-known how state transistion systems can be defined in Iris via Iris's notion of monoids and ghost state.)

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Capability machines are promising targets for secure compilers since capabilities can be used to enforce abstractions that are usually expected for high-level languages, such as well-bracketed control-flow (WBCF) and local state encapsulation (LSE).
  We present the first formalization of a capability machine that supports mechanized reasoning about deep semantic properties, including WBCF and LSE.
  Our formalization is done in the Coq implementation of Iris, a state-of-the-art concurrent higher-order separation logic, and includes a formalization of the logical relation defined by Skorstensgaard et al. \cite{skorstengaardESOP18}, which can used to prove WBCF and LSE.
\end{abstract}

@@

While Iris is a framework and supports many languages, it is geared towards models of higher-level languages, which abstract from the fact that programs are stored in memory, and hence come equipped with notions of expressions and values, in addition to the program memory.
In contrast, our low-level capability machine model has no notion of expression and values, it just consists of a memory and a register file.
The program counter register contains a pointer to an address in memory.
That address in turn will contain an integer, which can then be decoded to a machine instruction, such as Load, Store, Jump, etc.
Once an instruction has been executed, the program counter is updated and will then point to the next instruction in memory.

@@

% Similar to \cite{Jensen:2013}, the program logic rules for each instruction are of the following form: 
The Hoare triples for basic instructions take the following form
\begin{align*}
\small
	\textit{run time conditions } \wedge &\textit{ decode}(w) = \emph{instr} \Rightarrow  \\
	\{\{\{ \textsf{PC} \mapsto_r ((p,g),b,e,a) &* a \mapsto_a[p] w * ... \}\}\} \\ 
	\textsf{Instr Executable }\\
	\{\{\{\textsf{PC} \mapsto_r ((p,g),b,e,a&+1) * a \mapsto_a[p] w * ... \}\}\}
\end{align*} 
Here, the run time conditions correspond to the dynamic checks done by the capability machine, and \textsf{Instr Executable} is the abstract expression for executing the next instruction in memory.
This form of Hoare triple is similar to the one used in \cite{Jensen:2013} but
unlike \cite{Jensen:2013}, the decoding function is in our case assumed.

@@

\newcommand{\World}{\mathrm{World}} \newcommand{\Word}{\mathrm{Word}} We now outline how we define a unary logical relation in Iris that captures capability safety.
We define a value relation $\mathcal{V}$ as an Iris relation of type $\World \to \Word \to \emph{iProp}$, where $\World$ is a collection of state transition systems used to reason about local state.
(It is well-known how state transistion systems can be defined in Iris via Iris's notion of monoids and ghost state.)
